---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #1

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

Sofi's Answers:
```{r}
# Step 1: inspect the first few lines of the data set
library(classdata)
data("ames")
head(ames)
str(ames)
summary(ames)
```
Variables:
- Parcel ID (Character): The unique identifier for each property.
- Address (Character): The address of the property, including street name and city.
- Style (Factor): The architectural style of the property, with 12 levels (e.g., "1 Story Frame", "2 Story Frame").
- Occupancy (Factor): The type of occupancy for the property, with 5 levels (e.g., "Condominium", "Single-Family / Owner Occupied").
- Sale Date (Date): The date when the property was sold.
- Sale Price (Numeric): The sale price of the property. Range is from 0 to $20,500,000.
- Multi Sale (Character): Indicator of whether the property was part of a multiple sale. Most values are missing (NA).
- Year Built (Numeric): The year the property was built, ranging from 0 to 2022.
- Acres (Numeric): The size of the property in acres. Range from 0 to 12.012 acres.
- Total Living Area (sf) (Numeric): The total living area of the property in square feet. Range from 0 to 6007 square feet.
- Bedrooms (Numeric): The number of bedrooms in the property. Range from 0 to 10 bedrooms.
- Finished Basement Area (sf) (Numeric): The size of the finished basement in square feet. Range from 10 to 6496 square feet.
- Lot Area (sf) (Numeric): The size of the lot in square feet. Range from 0 to 523,228 square feet.
- AC (Character): Whether the property has air conditioning ("Yes"/"No").
- Fireplace (Character): Whether the property has a fireplace ("Yes"/"No").
- Neighborhood (Factor): The neighborhood the property is located in. There are 42 levels, each representing a different neighborhood.
Data Range Expectations:
- Sale Price: Expected to range from $0 (properties sold for $0) to a high value of $20,500,000.
- Year Built: The properties are built between 0 (indicating possibly missing data or incomplete records) to 2022.
- Acres: The property size ranges from 0 to 12.012 acres, with most properties having smaller lot sizes.
- Total Living Area (sf): The living area ranges from 0 to 6007 square feet. Properties with 0 square feet may indicate missing or incomplete data.
- Bedrooms: Properties range from 0 to 10 bedrooms, with a majority likely having 2-4 bedrooms.
- Finished Basement Area (sf): This ranges from 10 to 6496 square feet, with many properties lacking basement information (NA values).
- Lot Area (sf): The lot sizes range from 0 to 523,228 square feet. Smaller lots are expected, but there are outliers with very large lot sizes.
- Neighborhood: There are 42 neighborhoods, and the properties are distributed across these areas.

```{r}
#Step 4: answering pick a variable that might be related to the main variable. what is the range of that variable? plot. describe the pattern. Main variable is Sale Price
range(ames$`Sale Price`, na.rm = TRUE)

ggplot(ames, aes(x = `Sale Price`)) +
  geom_histogram(binwidth = 50000, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Sale Price", x = "Sale Price", y = "Count") +
  theme_minimal()

```
Pattern Description:
Right-Skewed Distribution:
- Most of the sale prices are concentrated on the lower end.
A long tail extends towards higher prices, indicating a few extremely expensive properties.
Extreme Outliers:
- A small number of properties have very high prices (above $5,000,000 and even close to $20,000,000).
These high prices likely represent commercial properties or luxury estates.

Tanish's Answers:
library(classdata)
data(ames)
dim(ames)            
summary(ames)

hist(ames$SalePrice,
     main = "Distribution of Sale Prices in Ames",
     xlab = "Sale Price",
     breaks = 30)
     
plot(ames$GrLivArea, ames$SalePrice,
     xlab = "Above Grade Living Area (sq ft)",
     ylab = "Sale Price ($)",
     main = "Sale Price vs. GrLivArea")
cor(ames$GrLivArea, ames$SalePrice, use = "complete.obs")

boxplot(SalePrice ~ OverallQual, 
        data = ames,
        xlab = "Overall Quality",
        ylab = "Sale Price ($)",
        main = "Sale Price by Overall Quality")

1. The first rows show variables like SalePrice, GrLivArea, OverallQual, and Neighborhood. SalePrice is in dollars, GrLivArea in square feet, OverallQual on a quality scale, and Neighborhood as discrete labels.

2. The main focus is SalePrice, which measures property value and is central to any home-sale analysis.

3. SalePrice typically ranges from around $30,000 to over $700,000. A histogram is right-skewed, with most homes clustered in the lower-to-mid range and a few high-priced outliers.

4. GrLivArea, ranging from a few hundred to over 3,000 square feet, shows a strong positive relationship with SalePrice. A scatterplot highlights that larger living areas generally command higher prices, with occasional outliers reflecting luxury properties.

